# Step 1
### Loading and preprocessing the data
Code for reading in the dataset and/or processing the data
```{r, echo = TRUE}
library(ggplot2)
activity <- read.csv("activity.csv")
```
Exploring the basics of this data
```{r}
dim(activity)
names(activity)
head(activity)
summary(activity)
```
# Step 2
### Histogram of the total number of steps taken each day
```{r, echo = TRUE}
activity_total_steps <- with(activity, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
names(activity_total_steps) <- c("date", "steps")
```
```{r fig.width=7, fig.height=6}
hist(activity_total_steps$steps, 
     main = "Total number of steps taken per day", 
     xlab = "Total steps taken per day",
     ylim = c(0,20), 
     breaks = seq(0,25000, by=2500))
```                          
# Step 3                       
### What is mean total number of steps taken per day?
### Mean and median number of steps taken each day
```{r, echo = TRUE}
mean(activity_total_steps$steps)
median(activity_total_steps$steps)
```                        
# Step 4                           
### What is the average daily activity pattern?
### Time series plot of the average number of steps taken
```{r, echo = TRUE}
average_daily_activity <- aggregate(activity$steps, by=list(activity$interval), FUN=mean, na.rm=TRUE)
names(average_daily_activity) <- c("interval", "mean")

```
```{r fig.width=7, fig.height=6}
plot(average_daily_activity$interval, 
     average_daily_activity$mean, 
     type = "l",
     xlab="Interval", 
     ylab="Average number of steps", 
     main="Average number of steps per intervals")
```
# Step 5
### Code to describe and show a strategy for imputing missing data
```{r, echo = TRUE}
sum(is.na(activity$steps))
imputed_steps <- average_daily_activity$mean[match(activity$interval, average_daily_activity$interval)]
activity_imputed <- transform(activity, steps = ifelse(is.na(activity$steps), yes = imputed_steps, no = activity$steps))
total_steps_imputed <- aggregate(steps ~ date, activity_imputed, sum)
names(total_steps_imputed) <- c("date", "daily_steps")
```                          
# Step 6
### Histogram of the total number of steps taken each day after missing values are imputed       
```{r fig.width=7, fig.height=6}                       
hist(total_steps_imputed$daily_steps, 
      xlab = "Total steps per day", 
      ylim = c(0,30), 
      main = "Total number of steps taken each day", 
     breaks = seq(0,25000,by=2500))
``` 
      
# Step 7
### Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends                      
```{r fig.width=7, fig.height=6}  
activity$date <- as.Date(strptime(activity$date, format="%Y-%m-%d"))
activity$datetype <- sapply(activity$date, function(x) {
  if (weekdays(x) == "sabato" | weekdays(x) =="domenica") 
  {y <- "Weekend"} else 
  {y <- "Weekday"}
  y
})
```  
    
# Step 8
### All of the R code needed to reproduce the results (numbers, plots, etc.) in the report
```{r, echo = TRUE}  
activity_by_date <- aggregate(steps~interval + datetype, activity, mean, na.rm = TRUE)
``` 
```{r fig.width=12, fig.height=6}              
ggplot(activity_by_date, aes(x = interval , y = steps, color = datetype)) +
       geom_line() +
       labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
       facet_wrap(~datetype, ncol = 1, nrow=2)

```

    